02-antecedentes

### 2024-04-24 01-introduction.qmd
render paper

---
title: "01-introduction"
author: "Equipo EDUMER"
format: pdf
editor: visual
---

@@

The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019]. This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021a]. Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].

>>>>

---
author: "Equipo EDUMER"
format: pdf
editor: visual
---

@@

The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019]. This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021]. Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary [@batruch_belief_2022].

### 2024-04-24 02-antecedentes.qmd
render paper

---
title: "02-antecedentes"
author: "Equipo EDUMER"
format: pdf
editor: visual
---

@@

Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegelBeliefsInequalityAmericans1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegenerDominantIdeologiesVariation1995] and cultural schemas [@homanBeingBecomingPoor2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castilloLegitimacyInequalityHighly2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education. The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).

>>>>

---
author: "Equipo EDUMER"
format: pdf
editor: visual
---

@@

Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; ] (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education. The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).
